---
title: "Current tournament predictions"
output: flexdashboard::flex_dashboard
author: "Mark Webster"
date: '`r format(Sys.time(), "Last modified: %d %b %Y")`'
---

```{r, "setup", include=FALSE}
current <- "CAMS20"
library(ggplot2)
theme_codex <- theme_dark() +
  theme(panel.border = element_rect(fill = NA, colour = "grey13"),
        plot.background = element_rect(fill = "grey13", colour = "grey13"),
        axis.ticks = element_line(colour = "white"),
        strip.background = element_rect(fill = "grey13", colour = NA),
        legend.background = element_rect(fill = "grey13", colour = NA),
        text = element_text(colour = "white"),
        axis.text = element_text(colour = "white"), 
  )
theme_set(theme_codex)
library(magrittr)
library(data.table)
library(codexdata)
library(codex)
library(DT)
library(lpSolve)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

```{r, "DT function to fill containers with single page"}
DT_fill <- function(df, dom_filter = FALSE) {
  datatable(df, rownames = FALSE, filter = "top", fillContainer = TRUE,
            options = list(dom = ifelse(dom_filter, "fti", "ti"), pageLength = nrow(df)))
}
```

```{r, "read model results"}
model_names <- c("Versus model, forum data",
                 "Versus model, forum + Metalize data")
results_file_subnames <- c("current_vs_split", "current_vs_split_metal")
results_file_names <- paste0("results/tidy_", results_file_subnames, ".rds")
vs_results <- stats::setNames(Map(readRDS, results_file_names), model_names)
```

# Entrants {data-navmenu="Data"}

```{r, "current matchups"}
current_entries <- decks[tournament == current,
                         .(player = factor(player, player),
                           deck = factor(deck, unique(deck)))]
kable(current_entries)
current_players <- as.character(current_entries$player)
current_decks <- as.character(current_entries$deck)
current_valid <- !is.na(current_decks)
```

# Completed matches {data-navmenu="Data"}

```{r, "current matches"}
current_matches <- matches[tournament == current]
```

```{r, "finished current matches"}
finished_current_matches <- current_matches[!is.na(victor)]
DT_fill(finished_current_matches[, c("round", "player1", "player2", "deck1", "deck2", "victor")],
        dom_filter = TRUE)
```

# Matchup plots: forum data {data-orientation=rows data-navmenu="Prediction"}

```{r, "matchup plot with wrap function"}
plot_samples_wrap <- function(matchup_samples) {
  plot_matchup_samples(matchup_samples[, .(P1 = `levels<-`(P1, stringr::str_replace_all(levels(P1), "/", " ")),
                                           P2 = `levels<-`(P2, stringr::str_replace_all(levels(P2), "/", " ")),
                                           prob_matchup, prob_mean)], 2) +
    facet_grid(P1~P2, labeller = labeller(P1 = label_wrap_gen(10), P2 = label_wrap_gen(15)))
}
```

### Deck only

```{r, "current forum data deck matchup samples plot", fig.width=10.5, fig.height=7.5, warning=FALSE}
plot_samples_wrap(get_matchups(vs_results[[1]],
                               as.character(current_decks[current_valid])))
```

### Overall

```{r, "current forum data overall matchup samples plot", fig.width=10.5, fig.height=7.5, warning=FALSE}
plot_matchup_samples(get_matchups(vs_results[[1]],
                                  current_decks[current_valid],
                                  current_players[current_valid]),
                     2)
```

# Matchup plots: all data {data-orientation=rows data-navmenu="Prediction"}

### Deck only

```{r, "current Metal deck matchup samples plot", fig.width=10.5, fig.height=7.5, warning=FALSE}
plot_samples_wrap(get_matchups(vs_results[[2]],
                               current_decks[current_valid]))
```

### Overall

```{r, "current Metal overall matchup samples plot", fig.width=10.5, fig.height=7.5, warning=FALSE}
plot_matchup_samples(get_matchups(vs_results[[2]],
                                  current_decks[current_valid],
                                  current_players[current_valid]),
                     2)
```

# Matchup tables: forum data {data-orientation=rows data-navmenu="Prediction"}

```{r, "current mean matchups"}
current_deck_matchup_array <- get_matchup_array(vs_results[[1]],
                                                current_decks[current_valid])
current_matchup_array <- get_matchup_array(vs_results[[1]],
                                           current_decks[current_valid],
                                           current_players[current_valid])

current_deck_matchups <- apply(current_deck_matchup_array, 2:3, mean)
current_matchups <- apply(current_matchup_array, 2:3, mean)
```

```{r, "current mean matchups with Metal data"}
current_metal_deck_matchup_array <- get_matchup_array(vs_results[[2]],
                                                      current_decks[current_valid])
current_metal_matchup_array <- get_matchup_array(vs_results[[2]],
                                                 current_decks[current_valid],
                                                 current_players[current_valid])

current_metal_deck_matchups <- apply(current_metal_deck_matchup_array, 2:3, mean)
current_metal_matchups <- apply(current_metal_matchup_array, 2:3, mean)
```

### Deck-only

```{r, "current forum data deck matchups"}
formatRound(DT_fill(reformat_mean_matchup(current_deck_matchups, "D"), dom_filter = TRUE), 3:4, 3L)
```

### Overall

```{r, "current forum data overall matchups"}
formatRound(DT_fill(reformat_mean_matchup(current_matchups, "P"), dom_filter = TRUE), 3:4, 3L)
```

# Matchup tables: all data {data-orientation=rows data-navmenu="Prediction"}

### Deck-only

```{r, "current all data deck matchups"}
formatRound(DT_fill(reformat_mean_matchup(current_metal_deck_matchups, "D"), dom_filter = TRUE), 3:4, 3L)
```

### Overall

```{r, "current all data overall matchups"}
formatRound(DT_fill(reformat_mean_matchup(current_metal_matchups, "P"), dom_filter = TRUE), 3:4, 3L)
```

# Nash equilibria {data-orientation=rows data-navmenu="Prediction"}

Nash equilibrium isn't really a good indicator of who's got the best chance, but it's a quick and dirty heuristic I can already do. Weights given are averaged over Nash equilibria calculated for each simulation sample (n = 4000).

```{r, "current mean Nash"}
current_deck_nash_samples <- get_nash_equilibria(current_deck_matchup_array)
current_deck_mean_nash <- as.data.table(apply(current_deck_nash_samples, 1:2, mean), 
                                        keep.rownames = "Player")[, c(.(Player = factor(Player, 
                                                                                        c("P1", "P2", "Both"))), 
                                                                      .SD), 
                                                                  .SDcols = dimnames(current_deck_nash_samples)[[2]]]
current_nash_samples <- get_nash_equilibria(current_matchup_array)
current_mean_nash <- as.data.table(apply(current_nash_samples, 1:2, mean), 
                                   keep.rownames = "Player")[, c(.(Player = factor(Player, 
                                                                                   c("P1", "P2", "Both"))), 
                                                                 .SD), 
                                                             .SDcols = dimnames(current_nash_samples)[[2]]]
```

```{r, "current mean Nash with all data"}
current_metal_deck_nash_samples <- get_nash_equilibria(current_metal_deck_matchup_array)
current_metal_deck_mean_nash <- as.data.table(apply(current_metal_deck_nash_samples, 1:2, mean), 
                                              keep.rownames = "Player")[, c(.(Player = factor(Player, 
                                                                                              c("P1", "P2", "Both"))), 
                                                                            .SD), 
                                                                        .SDcols = dimnames(current_metal_deck_nash_samples)[[2]]]
current_metal_nash_samples <- get_nash_equilibria(current_metal_matchup_array)
current_metal_mean_nash <- as.data.table(apply(current_metal_nash_samples, 1:2, mean), 
                                         keep.rownames = "Player")[, c(.(Player = factor(Player, 
                                                                                         c("P1", "P2", "Both"))), 
                                                                       .SD), 
                                                                   .SDcols = dimnames(current_metal_nash_samples)[[2]]]
```

## Mean Nash win rate

### Deck-only mean win rate, forum data

```{r, "forum data deck mean Nash win"}
kable(print_nash(current_deck_mean_nash))
```

### Overall mean win rate, forum data

```{r, "forum data overall mean Nash win"}
kable(print_nash(current_mean_nash))
```

### Deck-only mean win rate, all data

```{r, "all data deck mean Nash win"}
kable(print_nash(current_metal_deck_mean_nash))
```

### Overall mean win rate, all data

```{r, "all data overall mean Nash win"}
kable(print_nash(current_metal_mean_nash))
```

## Mean Nash pick rate

### Deck-only mean pick distribution, forum data

```{r, "forum data deck mean Nash spread"}
kable(reformat_used_nash(current_deck_mean_nash))
```

### Overall mean pick distribution, forum data

```{r, "forum data overall mean Nash spread"}
kable(reformat_used_nash(current_mean_nash))
```

### Deck-only mean pick distribution, all data

```{r, "all data deck mean Nash spread"}
kable(reformat_used_nash(current_metal_deck_mean_nash))
```

### Overall mean pick distribution, all data

```{r, "all data overall mean Nash spread"}
kable(reformat_used_nash(current_metal_mean_nash))
```

# Fairness {data-orientation=rows data-navmenu="Prediction"}

```{r, "current matchup progression"}
current_match_predictions <- current_matches[, .(round, round_match_number,
                                                 `Player 1` = player1, `Player 2` = player2,
                                                 `Deck 1` = deck1, `Deck 2` = deck2,
                                                 victor = factor(ifelse(is.na(victor), "ongoing",
                                                                        ifelse(victor == player1, "P1", "P2")),
                                                                 levels = c("P1", "P2", "ongoing")),
                                                 `forum data` = diag(current_matchups[player1, player2]),
                                                 `all data` = diag(current_metal_matchups[player1, player2]))] %>% 
  melt(measure.vars = c("forum data", "all data"), variable.name = "data", value.name = "P1 win probability")
current_match_predictions <- current_match_predictions[, c(.SD, .(fairness = 1 - 2*abs(`P1 win probability` - 1/2)))]
```

###

```{r, "plot current matchup progression"}
ggplot(current_match_predictions,
       aes(x = round, y = `P1 win probability`, colour = victor, shape = data)) +
  geom_point() +
  scale_colour_viridis_d() +
  ggtitle("Predicted matchup progression by tournament round")
```

###

```{r, "plot current fairness progression"}
ggplot(current_match_predictions,
       aes(x = round, y = `fairness`, colour = victor, shape = data)) +
  geom_point() +
  scale_colour_viridis_d() +
  ggtitle("Predicted fairness progression by tournament round")
```

# Prediction evaluation {data-orientation=rows data-navmenu="Prediction"}

```{r, "current counts"}
current_matches_result_tallies <- finished_current_matches[, .(P1 = as.factor(player1),
                                                               P2 = as.factor(player2),
                                                               `P1 win` = victor == player1)
                                                           ]
current_matches_counts <- current_matches_result_tallies[, .(wins = sum(`P1 win`),
                                                             losses = .N - sum(`P1 win`),
                                                             `P1 wins` = paste(sum(`P1 win`), .N, sep = "/")),
                                                         by = c("P1", "P2")]
```

```{r, "current prediction evaluation"}
current_matchups_table <- as.data.table(current_matchups,
                                        keep.rownames = "P1")[, c(list(P1 = factor(P1,
                                                                                   unique(P1))),
                                                                  .SD),
                                                              .SDcols = setdiff(colnames(current_matchups),
                                                                                "P1")] %>% 
  melt(id.vars = "P1", variable.name = "P2", value.name = "matchup")
current_forum_eval <- current_matchups_table[current_matches_counts, on = c("P1", "P2")
                                             ][, .(matchup, `P1 wins`,
                                                   `exp. score` = matchup*(1 - matchup),
                                                   `mean score` = weighted.mean(c(1 - matchup, matchup)^2,
                                                                                c(wins, losses))),
                                               by = c("P1", "P2")
                                               ][order(P1, P2)]
```

```{r, "CAWS19 prediction evaluation all data"}
current_metal_matchups_table <- as.data.table(current_metal_matchups,
                                              keep.rownames = "P1")[, c(list(P1 = factor(P1,
                                                                                         unique(P1))),
                                                                        .SD),
                                                                    .SDcols = setdiff(colnames(current_metal_matchups),
                                                                                      "P1")] %>% 
  melt(id.vars = "P1", variable.name = "P2", value.name = "matchup")
current_metal_eval <- current_metal_matchups_table[current_matches_counts, on = c("P1", "P2")
                                                   ][, .(matchup, `P1 wins`,
                                                         `exp. score` = matchup*(1 - matchup),
                                                         `mean score` = weighted.mean(c(1 - matchup, matchup)^2,
                                                                                      c(wins, losses))),
                                                     by = c("P1", "P2")
                                                     ][order(P1, P2)]
```

## Scores {data-height=200}

### Average Brier score

```{r, "current scores"}
kable(melt(current_matchups_table[current_metal_matchups_table, on = c("P1", "P2")
                                  ][, .(P1, P2, `forum data` = matchup, `all data` = i.matchup)
                                    ][current_matches_result_tallies, on = c("P1", "P2")
                                      ][, .(`forum data` = mean((`P1 win` - `forum data`)^2),
                                            `all data` = mean((`P1 win` - `all data`)^2))],
           id.var = character(0), variable.name = "data", value.name = "Brier score",
           variable.factor = FALSE)[, .(data = c("coin flip", data),
                                        `Brier score` = c(1/4, `Brier score`))])
```

## Match outcome DTs {data-height=800}

### forum data

```{r, "current forum data evaluation table"}
formatRound(DT_fill(current_forum_eval, dom_filter = TRUE), c(3, 5, 6), 3)
```

### all data

```{r, "current all data evalulation table"}
formatRound(DT_fill(current_metal_eval, dom_filter = TRUE), c(3, 5, 6), 3)
```
